---
title: "Optimization with two parameters in Gaussian distribution"
author: "Anderson Borba"
date: "18 June 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(ggthemes)
theme_set(theme_pander())
require(reshape2)
rm(list = ls())
require(maxLik)
library(plotly)
library(datasets)
#### webshot function
#webshot::install_phantomjs()
data(CO2)
set.seed(1234567890, kind="Mersenne-Twister")
```

# Example of Optimization with two parameters in Gaussian distribution

We assumed Gaussian distributions, so the densities are:
$$
f_{Z_(z; \mu, \sigma)} = \frac{1}{\sigma\sqrt{2\pi}}\exp\{-\frac{1}{2}(z-\mu)^2/\sigma^2\}
$$

We will illustrate in the following the idea of find parameters apply MLE (maximum Likelihood Estimation)

we will use log in both side the above equation.
$$
\ell(z; \mu, \sigma)=\log(f_{Z_(z; \mu, \sigma)}) = \log\Bigg(\frac{1}{\sigma\sqrt{2\pi}}\exp\{-(z-\mu)^2/2\sigma^2\}\Bigg)
$$
$$
\ell(z; \mu, \sigma)=\log\Bigg(\frac{1}{\sigma}\Bigg) + \log\Bigg(\frac{1}{\sqrt{2\pi}}\Bigg) + \log\Bigg(\exp\{-\frac{1}{2}(z-\mu)^2/\sigma^2\}\Bigg)
$$
$$
\ell(z; \mu, \sigma)=\log\big(\sigma^{-1}\big) + \log\big(\sqrt{2\pi}^{-1}\big)-\frac{1}{2\sigma^2}(z-\mu)^2
$$
$$
\ell(z; \mu, \sigma)=-\log\big(\sigma\big) - \log\big(\sqrt{2\pi}\big)-\frac{1}{2\sigma^2}(z-\mu)^2
$$
First, we obtain a strip of data from database $(CO_2)$ uptake of six plants from Quebec and six plants from Mississippi was measured at several levels of ambient $(CO_2)$ concentration. Half the plants of each type were chilled overnight before the experiment was conducted. More detail at [r documentation](https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/CO2)

The methods MLE is used to $(CO_2)$ database, where $n$ is the lenght of the database, and $z_i$ an element, 
$$
\mathcal{L}(\mu, \sigma)= \sum_{i=1}^n\ell(z_i; \mu, \sigma)
$$
$$
\mathcal{L}(\mu, \sigma)= \sum_{i=1}^n\Bigg[-\log\big(\sigma\big) - \log\big(\sqrt{2\pi}\big)-\frac{1}{2\sigma^2}(z_i-\mu)^2\Bigg]
$$
$$
\mathcal{L}(\mu, \sigma)= -\sum_{i=1}^n\log\big(\sigma\big) - \sum_{i=1}^n\log\big(\sqrt{2\pi}\big)-\frac{1}{2\sigma^2}\sum_{i=1}^n\Bigg[\frac{1}{2\sigma^2}(z_i-\mu)^2\Bigg]
$$
$$
\mathcal{L}(\mu, \sigma)= -n\log\big(\sigma\big) - n\log\big(\sqrt{2\pi}\big)-\frac{1}{2\sigma^2}\sum_{i=1}^n(z_i-\mu)^2
$$
```{r Define log-likelihood function}
n <- nrow(CO2)
z <- CO2$uptake
loglik <- function(theta) {
  mu <- theta[1]
  sigma <- theta[2]
  loglik <- -n*log(sqrt(2*pi)) - n*log(sigma) - 0.5*sum((z - mu)^2/sigma^2)
}
```
We can do the bidimentional plot for this function:
```{r Plot log-likelihood function}
n_mu <- 100
n_si <- 100
mu_i <- 1
mu_f <- 30
si_i <- 1
si_f <- 40
h_mu <- (mu_f - mu_i) / ( n_mu - 1)
h_si <- (si_f - si_i) / ( n_si - 1)
ne   <- n_mu * n_si
mesh_log_like <- matrix(0, n_mu, n_si)
eixo_mu <- rep(0,n_mu)
eixo_si <- rep(0, 1, n_si)
for (i in 1 : n_mu){
  eixo_mu[i] <- mu_i + (i - 1) * h_mu  
  for (j in 1 : n_si){
    eixo_si[j] <- si_i + (j - 1) * h_si
    mesh_log_like[i, j] <- loglik(c(eixo_mu[i], eixo_si[j]))
  }
}
p <- plot_ly(x = eixo_mu, y = eixo_si, z = mesh_log_like , type = "surface") 
p 
```

We apply optimization package to find the maximum point.
```{r Apply optimization}
res <- maxLik(loglik, start=c(mu=30, sigma=10), method = "BFGS")
res1 <- maxBFGS(loglik, start=c(mu=30, sigma=10))
pt  <- res$estimate
pt1 <- res1$estimate
```
<!-- Comentario-->